Recovery of a deleted SARS-CoV-2 deep sequencing data set from the early Wuhan epidemic

### 2021-06-15 paper/paper.tex
latest paper text updates

Numerous evolutionary analyses have shown that known human SARS-CoV-2 sequences are consistent with expansion from a single progenitor sequence~\citep{kumar2021evolutionary,pekar2021timing,rambaut2020dynamic,forster2020phylogenetic,pipes2021assessing}.
However, attempts to infer the identity of the SARS-CoV-2 progenitor sequence have been confounded by a perplexing fact: the sequences of the earliest reported isolates from Wuhan are \emph{not} the sequences most similar to SARS-CoV-2's bat coronavirus relatives~\citep{pipes2021assessing}.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/deltadist_RaTG13.pdf}
\caption{The reported collection date of each SARS-CoV-2 sequence versus its relative mutational distance from the RaTG13 bat coronavirus outgroup.
Mutational distances are relative to the putative progenitor proCoV2 inferred by \citet{kumar2021evolutionary}.
The plot shows all sequences in GISAID collected no later than February 28, 2020, after performing the quality-control described in the Methods.
Sequences that the joint WHO-China report~\cite{CITE} describes as being associated with the Wuhan Seafood Market are plotted with squares.
Points are slightly jittered on the y-axis.
Go to {\color{red} add link} for an interactive version of this plot that enables toggling of the outgroup to RpYN06 and RmYN02, mouseovers to see details for each point including strain name and mutations relative to proCoV2, and adjustment of the y-axis jittering.
Static versions of the plot with RpYN06 and RmYN02 outgroups are in Supplementary Figure~\ref{suppfig:deltadist_RpYN06_RmYN02}.
}
\label{fig:deltadist_RaTG13}
\end{figure*}

>>>>

Numerous evolutionary analyses have shown that known human SARS-CoV-2 sequences are consistent with expansion from a single progenitor sequence~\citep{kumar2021evolutionary,pekar2021timing,rambaut2020dynamic,forster2020phylogenetic,pipes2021assessing}.
However, attempts to infer the identity of the SARS-CoV-2 progenitor sequence have been confounded by a perplexing fact: the sequences of the earliest reported isolates from Wuhan are \emph{not} the sequences most similar to SARS-CoV-2's bat coronavirus relatives~\citep{pipes2021assessing}.
Although uncertainty remains about whether the proximal origin of SARS-CoV-2 was a zoonosis~\cite{CITE} or a lab accident~\cite{CITE}, all reasonable explanations agree that at a deeper level SARS-CoV-2 genome is largely derived from one or more bat coronaviruses~\cite{CITE}.
One would therefore expect that if SARS-CoV-2 is ultimately derived from a bat coronavirus
This fact is illustrated in Figure~\ref{fig:deltadist_RaTG13}.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/deltadist_RaTG13.pdf}
\caption{The reported collection date of each SARS-CoV-2 sequence versus its relative mutational distance from the RaTG13 bat coronavirus outgroup.
Mutational distances are relative to the putative progenitor proCoV2 inferred by \citet{kumar2021evolutionary}.
The plot shows all sequences in GISAID collected no later than February 28, 2020, after performing the quality-control described in the Methods.
Sequences that the joint WHO-China report~\citep{WHO2021origins} describes as being associated with the Wuhan Seafood Market are plotted with squares.
Points are slightly jittered on the y-axis.
Go to {\color{red} add link} for an interactive version of this plot that enables toggling of the outgroup to RpYN06 and RmYN02, mouseovers to see details for each point including strain name and mutations relative to proCoV2, and adjustment of the y-axis jittering.
Static versions of the plot with RpYN06 and RmYN02 outgroups are in Supplementary Figure~\ref{suppfig:deltadist_RpYN06_RmYN02}.
}
\label{fig:deltadist_RaTG13}
\end{figure*}
